---
title: "Center-Outward Quantile Regression"
output:
  html_document: default
  pdf_document: default
date: "2023-02-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a demonstration document showing how to implement Center-Outward Quantile Regression.

## Data generation
The following is the code for data generation:

```{r}
library(MASS)
library(parallel)
library(plot3D)
library(transport)
library(plotrix)
set.seed(Sys.time())

generatex<-function(n,d){ #$n$ is sample size and $d$ is the dimension of data
  x=runif(n*d,min = -1,max = 1)
  X=matrix(x,nrow = n,ncol = d)
  return(X)
}
generatey<-function(x){
  xsum=apply(abs(x), 1, sum)
  n=length(x[,1])
  Sigma <- matrix(c(1,0,0,1),2,2)
  e=mvrnorm(n=n, rep(0, 2), Sigma)
  xtemp=cbind(xsum^2,xsum^2)
  temp=(xsum+0.4)*e
  return(temp)
}
```

We set the parameters $n=3000,d=2 $ visualize the data

```{r}
n=3000
d=2
x= generatex(n,d)
y= generatey(x)    
X=matrix(x,nrow = n)
Y=matrix(y,nrow = n)
y1= y[,1]
y2= y[,2]
load(file='D:/testdata/mul_forest_x_3000_200.RData') #This is the R workspace when n=3000, m=2 and B=200.
scatter3D(x[,1], y1, y2,  type = "p",
          phi = 10, theta = 10,
          xlab='x',ylab='y1',zlab='y2',
          col = 'blue',pch =10, cex = 0.1,  
          ticktype = "detailed",bty = "b2")

```

This is the code to generate a uniform sphere
```{r}
generate_Grid<- function(nS,nR,n0){
  Y=matrix(NA,2,nS*nR)
  YY=array(NA,dim=c(2,nS,nR))
  for (i in 1:nR) {
    for (j in 1:nS) {
      YY[1,j,i]=i/(nR+1)*cos(2*pi*(j-1)/nS)
      YY[2,j,i]=i/(nR+1)*sin(2*pi*(j-1)/nS)
    }
    Y[,((i-1)*nS+1):(i*nS)] = YY[,,i]
  }
  Y0=matrix(0,2,n0)
  return(cbind(Y0,Y))
}
```

These are the code to create the empirical quantile countour
```{r}
opt_assign = function(data, grid){
  size = length(data[1, ])
  distMat = as.matrix(dist(rbind(t(data), t(grid))))
  distMatsub = distMat[1:size, (size+1):(2*size)]^2
  return(transport(rep(1, size), rep(1, size), costm = distMatsub, method = "networkflow")$to)
}

opt_qua = function(data,grid,weight){
  l_data = length(data[1,])
  l_grid = length(grid[1,])
  dismat = as.matrix(dist(rbind(t(grid),t(data))))
  subdist = dismat[1:l_grid,(l_grid+1):(l_data + l_grid) ]^2
  return( transport(rep(1/l_grid, l_grid), weight,  costm = subdist, method = "networkflow") )
}

#transport from weights to unit sphere

quantile_regions = function(tgrid,transdt,y){
  l=length(tgrid)
  out=matrix(rep(0,2*l),ncol=2)
  for (i in 1:l) {
    temp=transdt[ which(transdt$from==tgrid[i] ),]
    temp=temp[which.max(temp$mass),]
    if(length(temp$to)==1){
      out[i,]=y[temp$to,]
    }else{
      ytemp=y[temp$to,]
      yl=ytemp[,1]^2+ytemp[,2]^2
      out[i,]=ytemp[which.min(yl),]
    }
  }
  return(out)
}

```
These part is  generating  weights
```{r}
#forest weights
treeweight = function(tree , x ){
  index=1
  indexnow=1
  lr=1
  Threshold_value = tree[[1]]$Threshold_value
  feature=tree[[1]]$Feature_number
  while (!is.null( Threshold_value) ) {
    node=tree[[indexnow]]
    index=indexnow
    if(x[feature]<Threshold_value){
      indexnow = node[[5]][1]
      lr=1
    }else{
      indexnow = node[[5]][2]
      lr=2
    }
    Threshold_value = tree[[indexnow]]$Threshold_value
    feature=tree[[indexnow]]$Feature_number
  }
  if(lr==1){
    out=tree[[index]]$Idx_left
  }else{
    out=tree[[index]]$Idx_right
  }
  return(out)
}

forestweight = function(xtest,forest,ntree){
  lx=n
  ntree=length(forest)
  result=matrix(rep(0,ntree*lx),nrow = ntree )  
  for (i in 1:ntree) {
    tree=forest[[i]][[1]]
    index=forest[[i]][[2]]
    out=treeweight(tree,xtest)
    for (j in 1:length(out)) {
      result[i,index[out[j]]]=result[i,index[out[j]]]+1
    }
  }
  a=b=result
  weight1=rep(0,n)
  for (i in 1:length(a[,1])){
    b[i,]=a[i,]/sum(a[i,])
  }
  for (i in 1:length(X[,1])){
    weight1[i]=sum(b[,i])/ntree
  }
  return(weight1)
}


#kernel weights
kernels=function(x,b){
  #bandwidth
  x=x/b
  len=length(x[,1])
  gaussian=1:len
  for (i in 1:len) {
    gaussian[i]=1/(sqrt(2*pi))*exp(-sum(x[i,]^2)/2)
  }
  return(gaussian)
}
kernelweights=function(X,Xtest,b){
  Xtemp=X
  for (j in 1:length(X[,1])) {
    Xtemp[j,]=Xtemp[j,]- Xtest
  }
  K <-kernels(Xtemp,0.1)
  Ksum <- sum(K)
  weight1 <- K/Ksum
  return(weight1)
}


# KNN weights
euc.dist <- function(x1, x2) sqrt(sum((x1 - x2) ^ 2))

knnweight= function(xtest,X,N=50){
  cdist= rep(0,length(X[,1]))
  for (i in 1:length(X[,1])) {
    cdist[i]=euc.dist(xtest,X[i,])
  }
  rank=order(cdist)
  rank=rank[1:N]
  weight=rep(0,length(X[,1]))
  weight[rank]=1/N
  return(weight)
}
```

These part is  generating forest by parallel computation
```{r,eval= FALSE}
f<-function(k){
  library(MultivariateRandomForest)
  lx=length(X[,1])
  index=sample(1:lx,replace = T,lx)
  x=X[index,]
  y=Y[index,]
  x=matrix(x,nrow = lx)
  y=matrix(y,nrow = lx)
  Inv_Cov_Y = solve(cov(y))
  tree=build_single_tree(x, y, m_feature =d, min_leaf= 100,
                         Inv_Cov_Y= Inv_Cov_Y,Command = 2)
  return(list(tree,index))
}

ntree=200
out<-1:ntree
cl <- makeCluster(8)
clusterExport(cl,'X',envir = environment())
clusterExport(cl,'Y',envir = environment())
clusterExport(cl,'ntree',envir = environment())
clusterExport(cl,'d',envir = environment())
results <- parLapply(cl,out,f)
res <- do.call('list',results)
stopCluster(cl)
```

quantile contours

```{r}
N = 3001
NR = 30
NS = 100
N0 = N-NR*NS
Grid = generate_Grid(NS,NR,N0)

xt1=0.7
xt2=0.7
xt=1.4

weight1<-forestweight(c(xt1,xt2),res,ntree=ntree)

#weight1<- knnweight(cbind(xt1,xt2),X)

#weight1<-kernelweights(X,c(xt1 ,xt2),0.1)

df=opt_qua(t(y),Grid,weight1)
df=data.frame(df)

tau = 0.6
qt=quantile_regions((tau*NR*NS+N0+1):(tau*NR*NS+N0+NS),df,y)
qt1=qt[order(atan2(qt[,1]-mean(qt[,1]),qt[,2]-mean(qt[,2]))),]
qt2=rbind(qt1,qt1[1,])

tau =0.4
qqt=quantile_regions((tau*NR*NS+N0+1):(tau*NR*NS+N0+NS),df,y)
qqt1=qqt[order(atan2(qqt[,1]-mean(qqt[,1]),qqt[,2]-mean(qqt[,2]))),]
qqt2=rbind(qqt1,qqt1[1,])

tau = 0.2
qqqt=quantile_regions((tau*NR*NS+N0+1):(tau*NR*NS+N0+NS),df,y)
qqqt1=qqqt[order(atan2(qqqt[,1]-mean(qqqt[,1]),qqqt[,2]-mean(qqqt[,2]))),]
qqqt2=rbind(qqqt1,qqqt1[1,])

c=2.6
plot(qt2[,1],qt2[,2],type='l',asp=1,xlim = c(-c,c),ylim=c(-c,c),
     xlab = expression(Y[1]),ylab = expression(Y[2]),lwd=2)

lines(qqt2[,1],qqt2[,2],col=2,lwd=2,lty=1)
lines(qqqt2[,1],qqqt2[,2],col=3,lwd=2,lty=1)


draw.circle(0,0,1.333*(0.4+xt),lty=2,lwd = 2)
draw.circle(0,0,1.011*(0.4+xt),lty=2,border = 2,lwd = 2)
draw.circle(0,0,0.668*(0.4+xt),lty=2,border = 3,lwd = 2)
```
MSREC of $\tau=0.2,0.4,0.6$
```{r}

cat(mean(((qqqt[,1]^2+qqqt[,2]^2)^0.5 - (0.4+ xt)*0.668)^2  ),mean(((qqt[,1]^2+qqt[,2]^2)^0.5 - (0.4+ xt)*1.011)^2  )
    ,mean(((qt[,1]^2+qt[,2]^2)^0.5 - (0.4+ xt)*1.333)^2  ))

```
quantile tube, 
```{r,eval=FALSE}
l_test=20
x_test=seq (from = -0.9, to = 0.9, by = (1.8/ (l_test - 1 )))
xres=matrix(NA,nrow = l_test,ncol = NS+1)
yres=matrix(NA,nrow = l_test,ncol = NS+1)
zres=matrix(NA,nrow = l_test,ncol = NS+1)

for (i in 1:l_test) {
  weight1<-kernelweights(X,c(x_test[i],0.5),0.1)
  #weight1<-forestweight(c(x_test[i],0.5) ,res,ntree=ntree)
  #weight1<- knnweight(cbind(x_test[i],0.5),X,50)
  
  df=opt_qua(t(y),Grid,weight1)
  df=data.frame(df)
  tau = 0.2
  qt=quantile_regions((tau*NR*NS+N0+1):(tau*NR*NS+N0+NS),df,y)
  qt1=qt[order(atan2(qt[,1]-mean(qt[,1]),qt[,2]-mean(qt[,2]))),]
  qt2=rbind(qt1,qt1[1,])
  xres[i,]=t(qt2[,1])
  yres[i,]=t(qt2[,2])
  zres[i,]=rep(x_test[i],times=length(qt2[,1]))
}
mse=0
for (i in 1:length(xres[,1])){
  for (j in 1:length(xres[1,])){
    temp=(1/(abs(zres[i,j])+0.4+2))*((xres[i,j]-0)^2+(yres[i,j]-0)^2)^(0.5)
    mse=mse+abs( temp-0.668)^2
  }
}
print(mse/(length(xres[,1])*length(xres[1,]))) #This is the MSRET

```
We save the data and plot the tube in MATLAB
```{r}
# write.csv(t(xres),'D:/testdata/x_norm_02.csv', row.names = F)
# write.csv(t(yres),'D:/testdata/y_norm_02.csv', row.names = F)
# write.csv(t(zres),'D:/testdata/z_norm_02.csv', row.names = F)

```


